\documentclass[../notes.tex]{subfiles}

\pagestyle{main}
\renewcommand{\chaptermark}[1]{\markboth{\chaptername\ \thechapter\ (#1)}{}}
\setcounter{chapter}{6}

\begin{document}




\chapter[Group Action Applications: \texorpdfstring{$A_5$}{TEXT} and the Sylow Theorems]{Group Action Applications: \texorpdfstring{$\bm{A_5}$}{TEXT} and the Sylow Theorems}
\section[Actions of \texorpdfstring{$A_5$}{TEXT}]{Actions of \texorpdfstring{$\bm{A_5}$}{TEXT}}
\begin{itemize}
    \item \marginnote{11/7:}Classifying subgroups of $G=A_5\cong\text{Do}$.
    \item Let $H\leq G$. We must have $|H|\big||G|$ by Lagrange's theorem.
    \begin{itemize}
        \item Thus, if $H\leq A_5$, we must have
        \begin{equation*}
            |H| \in \{1,2,3,4,5,6,10,12,15,20,30,60\}
        \end{equation*}
    \end{itemize}
    \item A good place to start is with orders of $H$ that correspond to cyclic subsets.
    \item In particular, let's start with subgroups of the form $\gen{(**)(**)}$, which all have order 2.
    \begin{itemize}
        \item Are such groups conjugate?
        \item To prove that two groups of the form $\gen{(**)(**)}$ are conjugate, it will suffice to show that their generators are conjugate (since the only other element --- the identity --- will naturally be conjugate to itself).
        \item Let $x,y\in A_5$ be arbitrary elements of the form $(**)(**)$. Then there exists $g\in S_5$ such that $gxg^{-1}=y$.
        \item But is $g\in A_5$? If $g\in A_5$, then we are done. If $g\notin A_5$, then can we find an element $g'\in A_5$ such that $g'xg'^{-1}=y$?
        \item First, note that if $gxg^{-1}=y=g'xg'^{-1}$, then
        \begin{align*}
            g^{-1}(gxg^{-1})g' &= g^{-1}(g'xg'^{-1})g'\\
            x(g^{-1}g') &= (g^{-1}g')x
        \end{align*}
        Thus, $g^{-1}g'\in C_{S_5}(x)$, or $g'=gh$ for some $h\in C_{S_5}(x)$.
        \begin{itemize}
            \item If $g\notin A_5$ and we want $g'\in A_5$, then we must have $h\notin A_5$.
            \begin{itemize}
                \item Intuitively, this means that if $g$ is the product of an odd number of permutations and we want $g'=gh$ to be the product of an even number of permutations, $h$ had better be a product of an odd number of permutations as well.
                \item More formally, consider $G/A_5$. If $g\in gA_5\neq A_5$ and we want $g'\in g'A_5=A_5$, then by homomorphically mapping $gA_5$ to $1\in\Z/2\Z$ and $A_5$ to $0\in\Z/2\Z$, we must have $h\in gA_5$ to get $gh\in A_5$.
            \end{itemize}
            \item Regardless, this example motivates the following two propositions, which we can use to resolve the original conjugacy question.
        \end{itemize}
        \item By Proposition 1, since $x\sim y$ in $S_5$ and $C_{S_5}(x)\not\subset A_5$ (take the first transposition in $(**)(**)$; for example, know that $(12)$ commutes with $(12)(34)$), we know that $x\sim y$ in $A_5$.
        \item Therefore, there are 15 subgroups of the form $\gen{(**)(**)}$, all of which are conjugate in $A_5$.
    \end{itemize}
    \item Proposition 1: Let $x\sim y$ in $S_n$. Then if $C_{S_n}(x)\not\subset A_n$, then $x\sim y$ in $A_n$.
    \begin{proof}
        Since $x\sim y$ in $S_n$, there exists $g\in S_n$ such that $gxg^{-1}=y$. If $g\in A_n$, then we are done. Now suppose $g\notin A_n$. Since $C_{S_n}(x)\not\subset A_n$, there exists $h\in C_{S_n}(x)$ such that $hxh^{-1}=x$ and $h\notin A_n$. Since $g,h\notin A_n$, we have that $gh\in A_n$. Additionally, we have that
        \begin{equation*}
            (gh)x(gh)^{-1} = g(hxh^{-1})g^{-1}
            = gxg^{-1}
            = y
        \end{equation*}
        Therefore, $x\sim y$ in $A_n$, as desired.
    \end{proof}
    \item Proposition 2: If $C_{S_n}(x)\subset A_n$ and $\sigma x\sigma^{-1}=y$, then $x\sim y$ in $A_n$ iff $\sigma\in A_n$.
    \begin{proof}
        Suppose first that $x\sim y$ in $A_n$. Then $gxg^{-1}=y$ for some $g\in A_n$. Then as per the above, $gxg^{-1}=\sigma x\sigma^{-1}$ implies that $g^{-1}\sigma\in C_{S_n}(x)$. Thus, $\sigma=gh$ for some $h\in C_{S_n}(x)\subset A_n$. But since $g,h\in A_n$, we must have $\sigma\in A_n$, too.\par
        Now suppose that $\sigma\in A_n$. Then since $\sigma x\sigma^{-1}=y$, $x\sim y$ in $A_n$ as desired.
    \end{proof}
    \item Now we discuss subgroups of the form $\gen{(***)}$.
    \begin{itemize}
        \item Let $x$ be an arbitrary element of $A_5$ of the form $(***)$. In particular, suppose $x=(abc)$ for $a,b,c\in[5]$.
        \item Then $(de)\in C_{S_5}(x)$, where $d,e\in[5]$ are the other two elements that are not already represented by $a,b,c$.
        \item Moreover, $(de)$ will be in the centralizers of both $x$ and $x^2$.
        \item There are $\binom{5}{2}=10$ subgroups of the form we're discussing (20 generators/elements of the form $(***)$, though).
        \item Suppose we have two subgroups $\gen{x},\gen{y}$ of the form being discussed. We know that $\gen{x},\gen{y}$ are conjugate in $S_5$. But since $C_{S_5}(x)\not\subset A_5$ again as per the above, we know the groups are conjugate in $A_5$.
        \item Therefore, there are 10 subgroups of the form $\gen{(***)}$, all of which are conjugate in $A_5$.
    \end{itemize}
    \item Now we discuss subgroups of the form $\gen{(*****)}$.
    \begin{itemize}
        \item We know that $|C_{S_5}((12345))|\cdot|\{(12345)\}|=120$. Additionally, only a power of $(12345)$ commutes with it in this case, so the first term is 5. Thus, the second must be 24.
        \begin{itemize}
            \item In sum, we have showed that there are 24 elements conjugate to $(12345)$ in $S_5$.
            \item Another way we could show this is by counting all of the 5-cycles and knowing that they are all conjugate as 5-cycles. Indeed, there are $4!=24$ 5-cycles.
        \end{itemize}
        \item Claim: In $A_5$, $|x|=5$ implies $x\sim x$, $x\nsim x^2$, $x\nsim x^3$, and $x\sim x^4=x^{-1}$.
        \begin{proof}
            We know that $|x|=5$. Thus, let $x=(abcde)$.\par
            By the above statements on $C_{S_5}((12345))$, we know that $C_{S_5}(x)\subset A_5$. Thus, by proposition 2, $gxg^{-1}=x'$ iff $g\in A_n$. Thus,
            \begin{align*}
                exe^{-1} = x &\quad\Longrightarrow\quad x\sim x\\
                [(bc)(cd)(de)]x[(bc)(cd)(de)]^{-1} = (bced)(abcde)(bced)^{-1} = (acebd) &\quad\Longrightarrow\quad x\nsim x^2\\
                (bdec)(abcde)(bdec)^{-1} = (adbec) &\quad\Longrightarrow\quad x\nsim x^3\\
                [(be)(cd)](abcde)[(be)(cd)]^{-1} = (aedcb) &\quad\Longrightarrow\quad x\sim x^4=x^{-1}
            \end{align*}
            as desired.
        \end{proof}
        \item $x^2\sim x^3$ in $A_5$ as well.
        \item $(abced)$ and $(acebd)$ are conjugate by $(bce)\in A_5$.
        \item Six subgroups, all conjugate.
        \item All of the subgroups are conjugate, but not all of the elements are conjugate?
    \end{itemize}
    \item Consider $K=\{e,(12)(34),(13)(24),(14)(23)\}\triangleleft A_4\subset A_5$.
    \item Consider a transitive group action from $A_5$ to $X=\{\text{cong of }K\}$.
    \item $\Stab(K)=N_{A_5}(K)\supset A_4$.
    \item By O.S. trm, $X=|A_5|/|A_4|=5$.
    \item Let $H\subset A_5$ have $|H|=4$.
    \item We want to show that $H$ fixes a point. Equivalently, we want to find $x\in\{1,2,3,4,5\}$ such that $|\Orb(x)|=1$.
    \item Since $4=|H|=|\Orb(x)|\cdot|\Stab(x)|$ and $5\equiv 1\mod 2$. Thus, there is a fixed point.
    \item Thus, there are 15 cyclic subgroups of order 4 like $K$, and they are all conjugate.
    \item $H\leq A_5$ has index $d$ iff there is a transitive action and puts $A_5/H$. Induces a map from $A_5\to S_d$?? As $A_5$ has no normal subgroups. If $d=2,3,4$, ...?? If $d=5$, then $A_5\to S_5\to S_5/A_5$. But really $A_5\to S_5\to S_5/A_5\cong\Z/2\Z$.
    \item The hard ones are 6, 10, or 12.
    \item Consider a subgroup of $A_5$ of order 6. Must be $\Z/6\Z$ or $S_3$. These groups have subgroups of order 3. If we have this, it must be a subgroup of $S_3\times S_2\cap A_5$. Important: $\gen{(1,2,3)}$ and $(1,2)(4,5)$.
    \item Same analysis for subgroups of order 10. Subsets of order 1,2,5,10. $(12)$ orbits include...
    \item Table with sets.
    \item If we spend a couple of hours understanding this example in complete detail, that will be very helpful for the final.
\end{itemize}



\section{Blog Post: Actions of the Dodecahedral Group}
\emph{From \textcite{bib:Calegari}.}
\begin{itemize}
    \item \marginnote{11/26:}Recall that in HW2, we found a faithful action $\text{Do}\acts\text{5 inscribed cubes}$. This yielded an injective homomorphism $\text{Do}\to S_5$ identifying $\text{Do}$ with an order 60 subgroup. Moreover, this subgroup was necessarily $A_5$ since it is of order 60 and hence normal. Therefore,
    \begin{equation*}
        \text{Do} \cong A_5
    \end{equation*}
    \item Herein, we seek to classify all transitive actions of $\text{Do}$.
    \item \textbf{Equivalent} (group actions): Two group actions $G\acts X$ and $G\acts Y$ for which there exists a bijection $\phi:X\to Y$ satisfying
    \begin{equation*}
        \phi(g\cdot x) = g\cdot\phi(x)
    \end{equation*}
    for all $g\in G$ and $x\in X$.
    \item Because of the following theorem, to classify all transitive actions of $\text{Do}$, it will actually only be necessary to classify the conjugacy classes of the subgroups of $G$!
    \item Theorem: The transitive actions of a group $G$ up to equivalence are in bijection to the conjugacy classes of the subgroups of $G$.
    \begin{proof}
        To prove this claim, we will first define a map $f$ from the set of transitive actions of $G$ to the set of conjugacy classes of the subgroups of $G$, and a map $g$ from the set of conjugacy classes of the subgroups of $G$ to the set of transitive actions of $G$. We will then check that $f,g$ are well-defined, and that $g=f^{-1}$. Let's begin.\par
        Define\dots
        \begin{enumerate}
            \item $f$ by the rule, "take $X$ a set with a transitive action to the conjugacy class of $H=\Stab(x)$ for some $x\in X$;"
            \item $g$ by the rule, "take the conjugacy class of $H\leq G$ to $G\acts X=G/H$ by left multiplication."
        \end{enumerate}
        To prove that $f$ is well-defined, it will suffice to show that if $G\acts X$ and $G\acts Y$ transitive are equivalent, then $H=\Stab(x)$ for an arbitrary $x\in X$ and $H'=\Stab(y)$ for an arbitrary $y\in Y$ satisfy $H=\sigma H'\sigma^{-1}$ for some $\sigma\in G$. Suppose $G\acts X$ and $G\acts Y$ transitive are equivalent. Then there exists a bijection $\phi:X\to Y$ which preserves the group action. Let $H=\Stab(x)$ for some $x\in X$ arbitrary, and $H'=\Stab(y)$ for some $y\in Y$ arbitrary. Since $G\acts Y$ is transitive, $\phi(x)=\sigma\cdot y$ for some $\sigma\in G$. We choose this $\sigma$ to be our $\sigma$. To confirm that $H=\sigma H'\sigma^{-1}$, we will verify that $\sigma H'\sigma^{-1}\subset H$ and that $|\sigma H'\sigma^{-1}|=|H|$. Let $\sigma h'\sigma^{-1}\in\sigma H'\sigma^{-1}$ be arbitrary. Before we show that $\sigma h'\sigma^{-1}\cdot x=x$ (and hence $\sigma h'\sigma^{-1}\in\Stab(x)=H$), we prove one preliminary result. Indeed, we can show that like $\phi$, $\phi^{-1}$ also preserves the group action:
        \begin{align*}
            g\cdot\phi(x) &= \phi(g\cdot x)\\
            \phi^{-1}(g\cdot y) &= \phi^{-1}(\phi(g\cdot\phi^{-1}(y)))\\
            \phi^{-1}(g\cdot y) &= g\cdot\phi^{-1}(y)
        \end{align*}
        With this result, we have that
        \begin{align*}
            \sigma h'\sigma^{-1}\cdot x &= \sigma\cdot(h'\cdot(\sigma^{-1}\cdot x))\\
            &= \sigma\cdot(h'\cdot(\sigma^{-1}\cdot\phi^{-1}(\sigma\cdot y)))\\
            &= \sigma\cdot(h'\cdot\phi^{-1}(\sigma^{-1}\cdot(\sigma\cdot y)))\\
            &= \sigma\cdot(h'\cdot\phi^{-1}(y))\\
            &= \sigma\cdot\phi^{-1}(h'\cdot y)\\
            &= \sigma\cdot\phi^{-1}(y)\\
            &= \phi^{-1}(\sigma\cdot y)\\
            &= x
        \end{align*}
        as desired. As to the second statement we wish to verify, since $\phi$ is a bijection, $|X|=|Y|$. Thus, by the Orbit-Stabilizer theorem and the transitivity of both group actions,
        \begin{equation*}
            |H| = |\Stab(x)|
            = \frac{|G|}{|\Orb(x)|}
            = \frac{|G|}{|X|}
            = \frac{|G|}{|Y|}
            = \frac{|G|}{|\Orb(y)|}
            = |\Stab(y)|
            = |H'|
        \end{equation*}
        Since conjugate groups have the same order, $|H'|=|\sigma H'\sigma^{-1}|$. Therefore, by transitivity,
        \begin{equation*}
            |H| = |\sigma H'\sigma^{-1}|
        \end{equation*}
        as desired.\par
        To prove that $g$ is well-defined, it will suffice to show that $H\leq G$ and $\sigma H\sigma^{-1}\leq G$ map to equivalent transitive group actions. First off, since all actions of a group on its quotient groups are transitive as per the previous lecture, we know that we are mapping subgroups to \emph{transitive} group actions of $G$. Additionally, let $X=G/H$ and $Y=G/\sigma H\sigma^{-1}$. To confirm that $G\acts X$ and $G\acts Y$ are \emph{equivalent}, it will suffice to find a bijection $\phi:X\to Y$ that preserves the action. Define $\phi:X\to Y$ by
        \begin{equation*}
            \phi(\gamma H) = (\gamma\sigma^{-1})\sigma H\sigma^{-1} = \gamma H\sigma^{-1}
        \end{equation*}
        To confirm that $\phi$ is well-defined, it will suffice to verify that $\phi(\gamma H)=\phi(\gamma hH)$ for all $\gamma\in G$, $h\in H$. Let $\gamma\in G$, $h\in H$ be arbitrary. Then
        \begin{equation*}
            \phi(\gamma hH) = \gamma hH\sigma^{-1}
            = \gamma H\sigma^{-1}
            = \phi(\gamma H)
        \end{equation*}
        as desired.
        % We can also check that $\phi$ is a homomorphism and is bijective. 
        % To confirm that $\phi$ is a homomorphism, it will suffice to show that $\phi(\gamma H*\gamma' H)=\phi(\gamma H)\phi(\gamma' H)$ for all $\gamma,\gamma'\in G$. Let $\gamma,\gamma'\in G$ be arbitrary. Then
        % \begin{equation*}
        %     \phi(\gamma H*\gamma' H) = \phi(\gamma\gamma' H)
        %     = \gamma\gamma'\sigma^{-1}\sigma H\sigma^{-1}
        %     = \gamma\sigma^{-1}\sigma H\sigma^{-1}*\gamma'\sigma^{-1}\sigma H\sigma^{-1}
        %     = \phi(\gamma H)\phi(\gamma' H)
        % \end{equation*}
        % as desired.
        $\phi$ is naturally bijective since it takes as input $\gamma H$ for all $\gamma\in G$ (i.e., all cosets of $H$) and produces as output $(\gamma\sigma^{-1})\sigma H\sigma^{-1}$ (i.e., all cosets of $\sigma H\sigma^{-1}$ since all $\gamma\sigma^{-1}$'s are distinct by the Sudoku lemma).
        To confirm that $\phi$ preserves the group action, it will suffice to verify that $\phi(g\cdot\gamma H)=g\cdot\phi(\gamma H)$ for all $g\in G$ and $\gamma H\in X$. Let $g\in G$ and $\gamma H\in X$ be arbitrary. Then
        \begin{equation*}
            \phi(g\cdot\gamma H) = \phi(g\gamma H)
            = g\gamma H\sigma^{-1}
            = g\gamma\sigma^{-1}\sigma H\sigma^{-1}
            = g\cdot\gamma\sigma^{-1}\sigma H\sigma^{-1}
            = g\cdot\phi(\gamma H)
        \end{equation*}
        as desired.\par\smallskip
        To prove that $g=f^{-1}$, if will suffice to show that $f\circ g$ is the identity on the set of conjugacy classes of the subgroups of $G$ and $g\circ f$ is the identity on the set of transitive actions of $G$.\par
        Tackling $f\circ g$: Let $H\leq G$ be arbitrary. Then $g$ takes $H$ to the action of $G$ on $G/H$ by left multiplication, and $f$ takes $G/H$ back to $\Stab(\gamma H)$ for some $\gamma H\in G/H$. We now need only confirm that $\Stab(\gamma H)$ is conjugate to $H$. But since $\Stab(\gamma H)=\gamma H\gamma^{-1}$ by last lecture, we have the desired result.\par
        Tackling $g\circ f$: Let $X$ be an arbitrary set on which $G$ acts transitively. Then $f$ takes $X$ to the conjugacy class of $H=\Stab(x)$ for some $x\in X$, and $g$ takes $H$ back to the (transitive) action of $G$ on $G/H$ by left multiplication. To prove that these two actions are equivalent, it will suffice to find a bijection $\phi:G/H\to X$ that preserves the action. Define $\phi:G/H\to X$ by
        \begin{equation*}
            \phi(gH) = g\cdot x
        \end{equation*}
        where $x$ is the same element of $X$ used to define $H$. To confirm that $\phi$ is well-defined, it will suffice to verify that $\phi(gH)=\phi(ghH)$ for all $g\in G$, $h\in H$. Let $g\in G$, $h\in H$ be arbitrary. But since $h\in\Stab(x)$, we have that
        \begin{equation*}
            \phi(ghH) = gh\cdot x
            = g\cdot(h\cdot x)
            = g\cdot x
            = \phi(gH)
        \end{equation*}
        as desired. To confirm that $\phi$ is bijective, it will suffice to verify that $\phi$ is injective and surjective. For injectivity, we have that
        \begin{align*}
            \phi(gH) &= \phi(g'H)\\
            g\cdot x &= g'\cdot x
        \end{align*}
        so $g^{-1}g'\in\Stab(x)=H$. But this implies that $g'=gh$ for some $h\in H$, meaning that
        \begin{equation*}
            g'H = ghH = gH
        \end{equation*}
        as desired. For surjectivity, since $G\acts X$ is transitive, there exists $g\in G$ for which $g\cdot x=x'$ for all $x'\in X$. Therefore, for any $x'\in X$, $gH\in G/H$ satisfies
        \begin{equation*}
            \phi(gH) = g\cdot x = x'
        \end{equation*}
        as desired. To confirm that $\phi$ preserves the group action, it will suffice to verify that $\phi(\gamma\cdot gH)=\gamma\cdot\phi(gH)$ for all $\gamma\in G$ and $gH\in G/H$. Let $\gamma\in G$ and $gH\in G/H$ be arbitrary. Then
        \begin{equation*}
            \phi(\gamma\cdot gH) = \phi(\gamma gH)
            = \gamma g\cdot x
            = \gamma\cdot(g\cdot x)
            = \gamma\cdot\phi(gH)
        \end{equation*}
        as desired.
    \end{proof}
    \item Calegari reviews the isomorphism between $\text{Do}\cong\text{Ic}$.
    \item Subgroups of $A_5$ with distinct conjugacy classes.
    \begin{enumerate}
        \item The trivial subgroup.
        \item The cyclic group $\gen{(12)(34)}$ of order 2.
        \item The cyclic group $\gen{(123)}$ of order 3.
        \item The Klein 4-group $\gen{(12)(34),(13)(24),(14)(23)}$ of order 4.
        \item The cyclic group $\gen{(12345)}$ of order 5.
        \item The group $\gen{(123),(23)(45)}\cong S_3\cong D_6$ of order 6.
        \item The dihedral group $D_{10}=\gen{(12345),(25)(34)}$ of order 10.
        \item The group $A_4=\gen{(123),(124)}$ of order 12.
        \item The group $A_5$ of order 60.
    \end{enumerate}
    \item Notes on the above.
    \begin{itemize}
        \item These are actually \emph{all} of the subgroups of $A_5$.
        \item All of the above subgroups have different orders. Thus, there is a unique equivalence class of transitive actions on this group for a given set $X$ with
        \begin{equation*}
            |X| = 60,30,20,15,12,10,6,5,1
        \end{equation*}
    \end{itemize}
    \item Since $A_5$ has no non-trivial normal subgroups to act as kernels, all actions save the final one below will be faithful.
    \item Actions of the dodecahedral group:
    \begin{enumerate}
        \item The action on the "one" dodecahedron.
        \item The action on the five inscribed cubes.
        \item The action on the six pairs of opposite faces of the dodecahedron.
        \begin{enumerate}
            \item Equivalently, the action on the six pairs of opposite diagonals of the icosahedron.
        \end{enumerate}
        \item The action on the ten pairs of opposite vertices of the dodecahedron.
        \begin{enumerate}
            \item Equivalently, the action on the ten pairs of opposite faces of the icosahedron.
        \end{enumerate}
        \item The action on the twelve faces of the dodecahedron.
        \begin{enumerate}
            \item Equivalently, the action on the twelve vertices of the icosahedron.
        \end{enumerate}
        \item The action on the fifteen pairs of opposite edges of the dodecahedron.
        \begin{enumerate}
            \item Equivalently, the action on the fifteen pairs of opposite edges of the icosahedron.
        \end{enumerate}
        \item The action on the twenty vertices of the dodecahedron.
        \begin{enumerate}
            \item Equivalently, the action on the twenty faces of the icosahedron.
        \end{enumerate}
        \item The action on the thirty edges of the dodecahedron.
        \begin{itemize}
            \item Equivalently, the action on the thirty edges of the icosahedron.
        \end{itemize}
        \item The action of the group on itself by left multiplication.
    \end{enumerate}
\end{itemize}



\section[\texorpdfstring{$p$}{TEXT}-Groups]{\texorpdfstring{$\bm{p}$}{TEXT}-Groups}
\begin{itemize}
    \item \marginnote{11/9:}\textbf{$\bm{p}$-group}: A finite group of order $p^m$, where $p$ is prime and $m\geq 1$. \emph{Denoted by} $\bm{P}$.
    \item Example: If $|P|=p$, then $P\cong\Z/p\Z$.
    \item \textbf{Fixed point} (of $X$ under $G\acts X$): A point $x\in X$ for which $|\Orb(x)|=1$.
    \item Proposition: Let $P\acts X$ where $P$ is a $p$-group. Then the number of fixed points is congruent to $|X|\mod p$.
    \begin{proof}
        Let $x\in X$ be arbitrary. By the Orbit-Stabilizer theorem,
        \begin{equation*}
            p^m = |P| = |\Orb(x)|\cdot|\Stab(x)|
        \end{equation*}
        If $x$ is a fixed point, then $|\Orb(x)|=1$. However, if $x$ is not a fixed point, then we have by the above that no nontrivial element has order less than $p$ and hence $|\Orb(x)|\equiv 0\mod p$.\par
        As we know,
        \begin{equation*}
            X = \bigsqcup\text{Orbits}
            = \{\text{Fixed points}\}\sqcup\{\text{Non-trivial orbits}\}
        \end{equation*}
        Therefore, $|X|$ is equal to the number of fixed points plus the sum of the magnitudes of the other orbits. But since the magnitudes of the other orbits are all multiples of $p$ as per the above, we have that $|X|$ is congruent to the number of fixed points mod $p$. The desired result readily follows.
    \end{proof}
    \item Corollary: If $|X|\not\equiv 0\mod p$, then there exists at least one fixed point.
    \item \textbf{Center} (of $G$): The set of elements in $G$ that commute with every element of $G$. \emph{Denoted by} $\bm{Z(G)}$. \emph{Given by}
    \begin{equation*}
        Z(G) = \{g\in G\mid gx=xg\ \forall\ x\in G\}
    \end{equation*}
    \item Proposition: Let $P$ be a $p$-group, and $Z:=Z(P)$ be the center of $P$. Then $Z$ is a non-trivial normal subgroup.
    \begin{proof}
        To prove that $Z$ is normal, it will suffice to show that for all $x\in Z$ and $g\in G$, $gxg^{-1}\in Z$. Let $x\in Z$ and $g\in G$ be arbitrary. Then since $x\in Z$, $gx=xg$, i.e., $gxg^{-1}=x\in Z$, as desired.\par
        To prove that $Z$ is non-trivial, we make use of the previous proposition. Let $P\acts P$ by conjugation. We first prove that $Z(P)$ is exactly the set of fixed points of $P$. If $x\in P$ is a fixed point, then $pxp^{-1}=x$ for all $p$, so $x\in Z(P)$. In the other direction, if $x\in Z(P)$ normal, then by the definition of the center, $pxp^{-1}=x$ for all $p\in P$. Thus, $|Z(P)|$ is equal to the number of fixed points of $P$, and hence $|Z(P)|\equiv|P|\mod p\equiv 0\mod p$. Thus, we could have $|Z(P)|=0$, but since $e\in Z(P)$, we must instead have $|Z(p)|\geq p$. Therefore, $Z(P)$ is nontrivial.
    \end{proof}
    \item We get from this proposition an outline for "classifying" $p$-groups. We will do this inductively on $k$. Here are the steps.
    \begin{enumerate}
        \item Understand Abelian $p$-groups.
        \item Understand all $p$-groups of order $|p^k|$.
        \item Let $|P|=p^{k+1}$. Then by the above, $Z\triangleleft P$. If $Z=P$, use 1. If $Z\neq P$, then $|Z|$ and $|P/Z|$ divide $p^k$, so we can use 2.
    \end{enumerate}
    \item Goal: Knowing $Z$ and $G/Z$, try to find all possible $G$.
    \item Classification for $k=2$.
    \begin{enumerate}
        \item Abelian groups. By Lagrange's theorem, there are two possibilities: There exists $x$ with $|x|=p^2$, and there exists $x$ with $|x|=p$.
        \begin{enumerate}
            \item $G$ has an element of order $p^2$, and hence $G\cong\Z/p^2\Z$.
            \item There exists $x\in G$ such that $|x|=p$. Let $y\in G\setminus\gen{x}$. Then $y^p=e$. Thus, $G=\gen{x,y}$. $x^p=e=y^p$ and $xy=yx$. Thus, $G\cong\Z/p\Z\times\Z/p\Z$.
        \end{enumerate}
        \item Suppose $G$ is not abelian. $Z$ still has a nontrivial center, though, and hence any proper nontrivial subgroup of $G$ is necessarily isomorphic to $\Z/p\Z$ for the $k=2$ case. Thus, the only possible pair $(Z,G/Z)$ is $(Z,G/Z)=(\Z/p\Z,\Z/p\Z)$. But then $G/Z\cong\Z/p\Z$ is cyclic, so by HW4 Q5, $G$ is abelian, a contradiction. Therefore, $G\cong\Z/p^2\Z$ or $(\Z/p\Z)^2$, hence abelian.
    \end{enumerate}
    \item (Partial) classification for $k=3$.
    \begin{enumerate}
        \item Abelian groups: $\Z/p^3\Z$, $\Z/p^2\Z\times\Z/p\Z$, and $(\Z/p\Z)^3$.
        \item Possible pairs $(Z,G/Z)$:
        \begin{align*}
            (\Z_{p^2},\Z_p)^\times&&
            (\Z_p,\Z_{p^2})^\times&&
            (\Z_p^2,\Z_p)^\times&&
            (\Z_p,\Z_p^2)^\times
        \end{align*}
        $G/Z$ cyclic implies the same contradiction, so the only possibility is $Z=\Z_p$ and $G/Z=(\Z_p)^2$.
    \end{enumerate}
    \item Does the trend of no nonabelian groups continue for higher powers? No --- for $|G|=2^3=8$, both $D_8$ and $Q$ (the Quaternion group) are nonabelian counterexamples.
    \begin{itemize}
        \item Case 1: All elements in $G$ have order 2.
        \begin{itemize}
            \item $G$ is abelian: If $x,y\in G$ are arbitrary, then
            \begin{equation*}
                xy = xey
                = x(xy)^2y
                = xxyxyy
                = x^2yxy^2
                = eyxe
                = yx
            \end{equation*}
        \end{itemize}
        \item There are, of course, the other abelian groups as well. We now focus on the other case, and specifically its nonabelian forms.
        \item Case 2: There exists $g\in G$ with $|g|=4$.
        \begin{itemize}
            \item $g^2\neq e$.
            \item We also assume that $G$ is not abelian.
            \item $[G:\gen{g}]=2$, so $\gen{g}\triangleleft G$.
            \item Let $h\in G\setminus\gen{g}$. If $|h|=8$, then $G\cong\Z/8\Z$. But $G$ is not abelian, so this cannot be the case.
            \item Hence $|h|=2$ or $|h|=4$.
            \item If $|h|=4$, then $h^2\notin\gen{g}$ implies $G/\gen{g}\cong\Z/2\Z$ (another abelian case we are not interested in). Similarly, $h^2\in\gen{g}$ implies $h^2=g^2$. Thus, either $h^2=e$ or $h^2=g^2$.
            \item Since $\gen{g}\triangleleft G$, $hgh^{-1}\in\gen{g}$. It follows since the powers of $hgh^{-1}$ are as distinct as the powers of $g$ that $\gen{g}=\gen{hgh^{-1}}$. Thus, we either have $hgh^{-1}=g$ or $hgh^{-1}=g^{-1}$. In the first case, $hg=gh$, so $G=\gen{g,h}$ is abelian, and we are not interested.
            \item If $g^4=e=h^4$, then $G=Q$ and $hg=g^{-1}h$.
            \item If $g^4=e=h^2$, then $G=D_8$ and $hg=g^{-1}h$.
        \end{itemize}
    \end{itemize}
    \item We now investigate the case where $p$ is odd and $G=p^3$. Let $Z=\Z/p\Z$ and $G/Z=(\Z/p\Z)^2$.
    \begin{itemize}
        \item Consider a surjection $G\twoheadrightarrow G/Z$. Choose $x\mapsto(1,0)$ and $y\mapsto(0,1)$.
        \item Let $x^p,y^p,xyx^{-1}y^{-1}\in Z$.
        \item If $xy=yx$, then $G=\gen{x,y,Z}$ is abelian.
        \item Suppose $xy=yxz$ for some $z\in Z$ nontrivial.
        \item Case 1: All $g\in G$ have order $p$. Then
        \begin{equation*}
            G = \{y^bx^az^c\mid 0\leq a,b,c\leq p-1\}
        \end{equation*}
        \item We have that
        \begin{equation*}
            y^bx^az^c(y^Bx^Az^C) = y^bx^ay^Bx^Az^{c+C}
            = y^{b+B}x^{a+A}z^{c+C+aB}
        \end{equation*}
        since $xy=yxz$??
        \item This gets into $\text{GL}_3(\F_p)$, the group of $3\times 3$ invertible matrices over the field of numbers 0 to $p$ under addition mod $p$. In particular,
        \begin{equation*}
            \begin{pmatrix}
                1 & a & c\\
                0 & 1 & b\\
                0 & 0 & 1\\
            \end{pmatrix}
            \begin{pmatrix}
                1 & A & C\\
                0 & 1 & B\\
                0 & 0 & 1\\
            \end{pmatrix}
            =
            \begin{pmatrix}
                1 & a+A & c+c+aB\\
                0 & 1 & b+B\\
                0 & 0 & 1\\
            \end{pmatrix}
        \end{equation*}
    \end{itemize}
    \item $p$-groups and their orders for different values of $p,m$.
    \begin{table}[h!]
        \centering
        \small
        \renewcommand{\arraystretch}{1.2}
        \begin{tabular}{c|c|c|c|c}
             & $p$ & $p^2$ & $p^3$ & $p^4$\\
            \hline
            2 & 1 & 2 & $3+2$ & 14\\
            3 & 1 & 2 & $3+2$ & 15\\
            5 & 1 & 2 & $3+2$ & 15\\
            7 & 1 & 2 & $3+2$ & 15\\
        \end{tabular}
        \caption{$|P|$ for various $p,m$ values.}
        \label{tab:pGroupOrder}
    \end{table}
    \item Another perspective.
    \begin{itemize}
        \item Consider $x^p=e=y^p$, $xy=yxz$, $z^p=e$, and $z\in Z(P)$.
        \item Then
        \begin{equation*}
            (xy)^p = y^px^pz^{1+\cdots+p}
            = z^{p(p+1)/2}
        \end{equation*}
        \item If $p$ is odd, then $z^{p(p+1)/2}=e$ implies $(xy)^p=e$ \emph{except} when $p=2$.
    \end{itemize}
\end{itemize}



\section[Blog Post: \texorpdfstring{$p$}{TEXT}-Groups]{\texorpdfstring{Blog Post: $\bm{p}$}{TEXT}-Groups}
\emph{From \textcite{bib:Calegari}.}
\begin{itemize}
    \item \marginnote{11/27:}Mostly review of lecture.
    \item Claim (by Lagrange): Any subgroup of a $p$-group $P$ is also a $p$-group. The order of any element of $P$ is a power of $p$.
    \item \textbf{Fixed points} (of $X$ under $G\acts X$): The set of all fixed points of $X$. \emph{Denoted by} $\bm{\Fixed(X)}$.
    \item Theorem:
    \begin{equation*}
        |\Fixed(X)| \equiv |X|\mod p
    \end{equation*}
    \item Example: Let $X=P$ and $P\acts P$ by left multiplication. Then $|P|\mod p=p^m\mod p=0\mod p$. This squares with the fact that by the Sudoku lemma, $P$ has no fixed points ($|P|>1$ by definition since the smallest prime number is 2), so $|\Fixed(P)|=0\equiv 0\mod p$.
    \item Following up on the center being a non-trivial normal subgroup, we have the following corollary.
    \item More on $\text{GL}_3(\F_p)$.
    \begin{itemize}
        \item Let $P\subset\text{GL}_3(\F_p)$ be the set of matrices of the following form.
        \begin{equation*}
            \begin{pmatrix}
                1 & * & *\\
                0 & 1 & *\\
                0 & 0 & 1\\
            \end{pmatrix}
        \end{equation*}
        \item $P$ is non-abelian since the operation is matrix multiplication, which is not commutative.
        \item $|P|=p^3$ since we have 3 free entries in the matrix, each of which can take on $p$ values.
        \item Every matrix in $P$ is invertible (and hence an element of $\text{GL}_3(\F_p)$) since the determinant for such an upper triangular matrix is the product of the diagonal entries and hence $1\neq 0$.
        \item $P$ is closed under inversion since
        \begin{equation*}
            \begin{pmatrix}
                1 & x & y\\
                0 & 1 & z\\
                0 & 0 & 1\\
            \end{pmatrix}^{-1}
            =
            \begin{pmatrix}
                1 & -x & -y+xz\\
                0 & 1 & -z\\
                0 & 0 & 1\\
            \end{pmatrix}
        \end{equation*}
        \item $P$ is closed under multiplication since
        \begin{equation*}
            \begin{pmatrix}
                1 & a & b\\
                0 & 1 & c\\
                0 & 0 & 1\\
            \end{pmatrix}
            \begin{pmatrix}
                1 & x & y\\
                0 & 1 & z\\
                0 & 0 & 1\\
            \end{pmatrix}
            =
            \begin{pmatrix}
                1 & a+x & b+y+az\\
                0 & 1 & c+z\\
                0 & 0 & 1\\
            \end{pmatrix}
        \end{equation*}
        \item Exercises:
        \begin{enumerate}
            \item If $p\geq 3$, then every non-trivial element of $P$ has order $p$. In particular, knowing that every element of a group $P$ satisfies $x^p=e$ does not imply that $P$ is abelian unless $p=2$, in which case we did prove that such a group is abelian.
            % \begin{proof}
            %     Let
            %     \begin{equation*}
            %         \begin{pmatrix}
            %             1 & a & b\\
            %             0 & 1 & c\\
            %             0 & 0 & 1\\
            %         \end{pmatrix}
            %     \end{equation*}
            %     be an arbitrary non-trivial element of $P$. Then
            %     \begin{equation*}
            %         \begin{pmatrix}
            %             1 & a & b\\
            %             0 & 1 & c\\
            %             0 & 0 & 1\\
            %         \end{pmatrix}^2
            %         =
            %         \begin{pmatrix}
            %             1 & 2a & 2b+ac\\
            %             0 & 1 & 2c\\
            %             0 & 0 & 1\\
            %         \end{pmatrix}
            %     \end{equation*}
            % \end{proof}
            \item What are the order four elements in $P$ when $p=2$?
            \item The center $Z(P)$ is the subgroup of order $p$ containing all elements of the form
            \begin{equation*}
                \begin{pmatrix}
                    1 & 0 & *\\
                    0 & 1 & 0\\
                    0 & 0 & 1\\
                \end{pmatrix}
            \end{equation*}
            Moreover, we have $P/Z=(\Z/p\Z)^2$. In fact, I claim that if $P$ is any non-abelian group of order $p^3$, then $P/Z=(\Z/p\Z)^2$ has to be true. Since $P$ is not abelian, we can't have $Z=P$. We also know since $P$ is a $p$-group that $Z$ is non-trivial. Finally, we can't have $|Z|=p^2$ since then $P/Z$ would be cyclic and then (by HW4 Q5 again) $P=Z$, which is a contradiction. Finally, even when $|Z|=p$, we can't have $P/Z\cong\Z/p^2\Z$ since that still implies by HW4 Q5 that $P$ is cyclic and thus $P=Z$, another contradiction.
            \item When $p=2$, is this group the dihedral group $D_8$ or the quaternion group $Q$? (It is one of these groups --- see the claim below.)
        \end{enumerate}
        \item Claim: There are exactly 5 groups of order $p^3$, three of which are abelian and two of which are not. When $p>2$, the non-abelian groups of order $p^3$ are distinguished by whether all elements in $P$ have order dividing $p$ or not. When $p=2$, the two non-abelian groups are $D_8$ and $Q$.
        \begin{proof}
            No proof given.
        \end{proof}
        \item Exercise: Can you construct the other non-abelian group of order $p^3$ which has an element of order $p^2$?
    \end{itemize}
\end{itemize}



\section{Sylow I-II}
\begin{itemize}
    \item \marginnote{11/11:}\textbf{$\bm{p}$-Sylow}\footnote{Sylow is pronounced "SIH-lohv."}: A subgroup $P\leq G$ of order $|P|=p^n$ for some prime $p$ and $n\in\N$, where $G$ is a finite group of order $|G|=p^n\cdot k$ for $\gcd(p,k)=1$.
    \item Theorem (Sylow I --- Existence): Let $G$ be a finite group with order divisible by $p$. Then $G$ has a $p$-Sylow subgroup.
    \begin{proof}
        Let $X$ be the set of all subsets (not subgroups!) of $G$ of order $p^n$. Define $G\acts X$ by left multiplication. Then if $S=\{s_1,\dots,s_{p^n}\}\in X$, we have for instance that
        \begin{equation*}
            g\cdot S = gS = \{gs_1,\dots,gs_{p^n}\}
        \end{equation*}
        We now investigate the properties of $\Stab(S)$; we will eventually prove that there exists an $S$ for which $\Stab(S)$ is the desired $p$-Sylow. Let's begin.\par
        We will first show that $|\Stab(S)|\leq p^n$. Pick a $g\in\Stab(S)$. By definition $gs_1\in S$, so $gs_1=s_i$ for some $i=1,\dots,p^n$. It follows that $g=s_is_1^{-1}$. Thus, every element of $\Stab(S)$ is of the form $s_is_1^{-1}$, so there are at most $p^n$ elements in the set (one for each $i$).\par
        We now divide into two cases ($|\Stab(S)|=p^n$ for some $S$ and $|\Stab(S)|<p^n$ for all $S$). In the former case, we may choose $P=\Stab(S)$ to be our $p$-Sylow, and we are done. In the latter case, we can derive a contradiction, meaning that the former case is always true. To do so, let $S\in X$ be arbitrary. Note that by the Orbit-Stabilizer theorem,
        \begin{equation*}
            |\Stab(S)|\cdot|\Orb(S)| = |G| = p^n\cdot k \equiv 0\mod p^n
        \end{equation*}
        Since $|\Stab(S)|<p^n$, we know that $|\Stab(S)|\not\equiv 0\mod p^n$. It follows that the largest power of $p$ dividing $|\Stab(S)|$ (which we will call $m$) is less than $n$ (note that it is possible that $m=0$). But since $|G|$ is divisible by $p^n$ and $|\Stab(S)|$ is not, we have that
        \begin{equation*}
            |\Orb(S)| = \frac{|G|}{|\Stab(S)|}
            = \frac{p^n\cdot k}{p^m\cdots}
            = p^{n-m}\cdots
        \end{equation*}
        i.e., that $|\Orb(S)|$ has at least one power of $p$ in its prime factorization. This implies that $|\Orb(S)|\equiv 0\mod p$. But since $|\Orb(S)|$ is divisible by $p$ for all $S$, $|X|$ must be, too (why??). However,
        \begin{equation*}
            |X| = \binom{p^nk}{p^n}
            = \frac{(p^nk)!}{(p^nk-p^n)!p^n!}
            = \frac{(p^nk)(p^nk-1)\cdots(p^nk-p^n+1)}{(p^n)(p^n-1)\cdots 1}
            = \frac{p^nk}{p^n}\cdots\frac{p^nk-(p^n-1)}{p^n-(p^n-1)}
        \end{equation*}
        We show that every power of $p$ in the numerator above cancels with one in the denominator. In fact, we can do this term-by-term. Consider $p^nk-i$ and $p^n-i$ for some $i=0,\dots,p^n-1$. Let $p^j$ be the largest power of $p$ dividing $i$. Note that since $i<p^n$, we must have $j<n$. Thus, $p^j$ will divide $p^nk$ and $p^n$, too, and hence the differences $p^nk-i$ and $p^n-i$ as well. This implies the desired result. Therefore, since there are no "excess" powers of $p$ in the numerator above, $|X|$ is \emph{not} divisible by $p$, a contradiction.
    \end{proof}
    \item Example: Let $G=S_p$.
    \begin{itemize}
        \item $|G|=p!=p\cdot k$.
        \item Need to find a subgroup of order $p$.
        \item $P=\gen{(1,2,\dots,p)}$ is a $p$-Sylow of $G$.
    \end{itemize}
    \item Example: Let $G=S_4$.
    \begin{itemize}
        \item Pick $p=2$ so that $|G|=24=2^3\cdot 3$.
        \item Need to find a subgroup of order 8.
        \item We can choose $D_8\leq S_4$.
    \end{itemize}
    \item Theorem (Sylow II --- Uniqueness up to conjugation): Fix $P$ a $p$-Sylow.
    \begin{enumerate}
        \item If $Q\subset G$ is a $p$-Sylow, then $Q=gPg^{-1}$ for some $g\in G$.
        \item If $Q\subset G$ is a $p$-group, then $Q\subset gPg^{-1}$ for some $g\in G$.
    \end{enumerate}
    \begin{proof}
        Ask in office hours??
    \end{proof}
\end{itemize}




\end{document}